Mechanised Modal Model Theory

### 2019-11-18 p-modalmodels/paper.tex
omit type parameter in p-modalmodels

\section{Syntax, Semantics and the Standard Translation}[1.5p]
In our formalization, we only consider the basic modal language, in which the only primitive modal operator is the `$\Diamond$'. For a type $\alpha$, an $\alpha$-modal formula is either of form \HOLinline{\HOLConst{VAR}\;\HOLFreeVar{p}}, where $p$ is of type $\alpha$, or a disjunction $\phi\lor \psi$ of two $\alpha$-modal formulas, or the falsity $\bot$, or a negation $\lnot \phi$ of an $\alpha$-modal formula $\phi$, or, finally, of the form $\Diamond \phi$ where $\phi$ is an $\alpha$-modal formula. We create a data type called `form' of the formulas of this modal language. 
\begin{defn}
{\upshape\cite[Definition 1.9]{Blackburn}}
\begin{holmath}
  \ensuremath{\alpha}\;\HOLTyOp{form}\;=\;\HOLConst{VAR}\;\ensuremath{\alpha}\;\HOLTokenBar{}\;\HOLConst{DISJ}\;(\ensuremath{\alpha}\;\HOLTyOp{form})\;(\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLTokenBar{}\;\HOLSymConst{\ensuremath{\bot}}\;\HOLTokenBar{}\;(\HOLSymConst{\HOLTokenNeg{}})\;(\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLTokenBar{}\;\HOLConst{\ensuremath{\Diamond}}\;(\ensuremath{\alpha}\;\HOLTyOp{form})
\end{holmath}
\end{defn}

@@

  (\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model} = \HOLTokenLeftrec{}\;\HOLFieldName{frame}\;{:}\;\ensuremath{\beta}\;\HOLTyOp{frame};\;\HOLFieldName{valt}\;{:}\;\ensuremath{\alpha}\;\ensuremath{\rightarrow}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}\;\HOLTokenRightrec{}

@@

  (\HOLFreeVar{\ensuremath{\phi\sb{1}}} :\ensuremath{\alpha}\;\HOLTyOp{form})\equiv\sb{({:}\ensuremath{\beta})}(\HOLFreeVar{\ensuremath{\phi\sb{2}}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLTokenDefEquality{}\\
\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model})\;(\HOLBoundVar{w} :\ensuremath{\beta}).\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{1}}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{2}}}\\[3mm]

@@

\section{Basic Results}%[2p]
We discuss some highlights of formalized results in section 2.1-2.3 in \cite{Blackburn} below.
\subsection{Tree-like property}
A tree-like model is a model whose underlying frame is a tree. If $H$ is a tree with root $r$, we write \HOLinline{\HOLConst{tree}\;\HOLFreeVar{H}\;\HOLFreeVar{r}} in HOL:
\begin{defn}
{\upshape\cite[Definition 1.7]{Blackburn}}
\begin{holmath}
   \HOLConst{tree}\;\HOLFreeVar{H}\;\HOLFreeVar{r}\;\HOLTokenDefEquality{}\\
\;\;\HOLFreeVar{r}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenImp{}}\;{\HOLFreeVar{H}.\HOLFieldName{rel}}\ensuremath{\mid}\sb{\HOLFreeVar{H}.\HOLFieldName{world}}\HOLSymConst{\HOLTokenSupStar{}}\;\HOLFreeVar{r}\;\HOLBoundVar{w})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenNeg{}}\HOLFreeVar{H}.\HOLFieldName{rel}\;\HOLBoundVar{w}\;\HOLFreeVar{r})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLFreeVar{r}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenUnique{}}\HOLBoundVar{w}\sb{\mathrm{0}}.\;\HOLBoundVar{w}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{H}.\HOLFieldName{rel}\;\HOLBoundVar{w}\sb{\mathrm{0}}\;\HOLBoundVar{w}
\end{holmath}
\end{defn}
The tree-like property says an satisfiable modal formula can be satisfied in a tree-like model. It is proved in HOL as:
\begin{prop}
{\upshape\cite[Proposition 2.15]{Blackburn}}\label{2.15}
\begin{holmath}
  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model}),(\HOLFreeVar{w} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\ensuremath{\alpha},\;\ensuremath{\beta}\;\HOLTyOp{list})\;\HOLTyOp{model})\;(\HOLBoundVar{r} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{tree}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{frame}\;\HOLBoundVar{r}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{r}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
\end{holmath}
\end{prop}
The world set of the tree like model constructed from \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} according to the standard proof of this theorem is a set of lists of worlds in \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}}, that is why passing to a tree-like model does not preserve the type.

@@

\subsection{Finite model property}
There are two classical approach of constructing finite models using model theory, namely via selection and via filtration. 
\subsubsection{Selection}
Given \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\sb{\mathrm{1}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}, where $\phi$ has degree $k$, we can construct \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}}, \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}}, \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}} and \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} comsecutively, such that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} is the finite model we want, where:
\begin{itemize}
\item \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} is the tree-like model obtained from \ref{2.15} with root \HOLinline{\HOLFreeVar{w}\sb{\mathrm{2}}} such that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLFreeVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}.
\item \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}} is the restriction of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} to height $k$.
\item \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}} is obtained from \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}} by modifying the valuation into \HOLinline{\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{v}.\;\HOLKeyword{if}\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLConst{prop_letters}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLKeyword{then}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}\HOLSymConst{\ensuremath{\sp{V}}}\;\HOLBoundVar{p}\;\HOLBoundVar{v}\;\HOLKeyword{else}\;\HOLConst{F}}.
\end{itemize}
The construction of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} requires a lemma: 
\begin{lm}
{\upshape\cite[Proposition 2.29]{Blackburn}}
\begin{holmath}
  \ensuremath{\HOLTokenTurnstile}\HOLConst{FINITE}\;(\HOLFreeVar{\ensuremath{\Phi}} :\ensuremath{\alpha}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{INFINITE}\;\ensuremath{{\cal{U}}}({:}\ensuremath{\beta})\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{n} :\HOLTyOp{num}).\;\HOLConst{FINITE}\;{\HOLTokenLeftbrace{}\HOLBoundVar{\ensuremath{\phi}}\;\HOLTokenBar{}\;\HOLConst{DEG}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLBoundVar{n}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{prop_letters}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\Phi}}\HOLTokenRightbrace{}} /{\equiv\sb{(:\beta)}}
\end{holmath}
\end{lm}
We require the assumption that the universe of $\beta$ is infinite since we relies on the fact that two modal formulas \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{1}}}} and \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{2}}}} are equivalent if and only if \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{1}}}} and \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{2}}}} are equivalent. This is easy to prove using set theory. However, in simple type theory, the proof of  \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{1}}}\equiv\sb{({:}\ensuremath{\beta})}\HOLFreeVar{\ensuremath{\phi\sb{2}}}} iff \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{1}}}\equiv\sb{({:}\ensuremath{\beta})}\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{2}}}} requires the infiniteness of the type universe of \HOLinline{\ensuremath{\beta}}.

@@

As we used \ref{2.15}, we change the type of the model by passing to a finite model via selection:
\begin{thm}
{\upshape\cite[Theorem 2.34 (Finite model property, via selection)]{Blackburn}}
\begin{holmath}
  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :(\ensuremath{\alpha},\;\ensuremath{\beta}\;\HOLTyOp{list})\;\HOLTyOp{model})\;(\HOLBoundVar{v} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{v}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
\end{holmath}
\end{thm}

>>>>

\section{Syntax, Semantics and the Standard Translation}[1.5p]
In our formalization, we only consider the basic modal language, in which the only primitive modal operator is the `$\Diamond$'. For a type $\alpha$, an $\alpha$-modal formula is either of form \HOLinline{\HOLConst{VAR}\;\HOLFreeVar{p}}, where $p$ is of type $\alpha$, or a disjunction $\phi\lor \psi$ of two $\alpha$-modal formulas, or the falsity $\bot$, or a negation $\lnot \phi$ of an $\alpha$-modal formula $\phi$, or, finally, of the form $\Diamond \phi$ where $\phi$ is an $\alpha$-modal formula. We create a data type called `form' of the formulas of this modal language. 
\begin{defn}
{\upshape\cite[Definition 1.9]{Blackburn}}
\begin{holmath}
  \HOLTyOp{form}\;=\;\HOLConst{VAR}\;\HOLTyOp{num}\;\HOLTokenBar{}\;\HOLConst{DISJ}\;\HOLTyOp{form}\;\HOLTyOp{form}\;\HOLTokenBar{}\;\HOLSymConst{\ensuremath{\bot}}\;\HOLTokenBar{}\;(\HOLSymConst{\HOLTokenNeg{}})\;\HOLTyOp{form}\;\HOLTokenBar{}\;\HOLConst{\ensuremath{\Diamond}}\;\HOLTyOp{form}
\end{holmath}
\end{defn}

@@

  \ensuremath{\beta}\;\HOLTyOp{model} = \HOLTokenLeftrec{}\;\HOLFieldName{frame}\;{:}\;\ensuremath{\beta}\;\HOLTyOp{frame};\;\HOLFieldName{valt}\;{:}\;\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\ensuremath{\beta}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool}\;\HOLTokenRightrec{}

@@

  (\HOLFreeVar{\ensuremath{\phi\sb{1}}} :\HOLTyOp{form})\equiv\sb{({:}\ensuremath{\beta})}(\HOLFreeVar{\ensuremath{\phi\sb{2}}} :\HOLTyOp{form})\;\HOLTokenDefEquality{}\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :\ensuremath{\beta}\;\HOLTyOp{model})\;(\HOLBoundVar{w} :\ensuremath{\beta}).\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{1}}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi\sb{2}}}\\[3mm]

@@

\section{Basic Results}%[2p]
We discuss some highlights of formalized results in section 2.1-2.3 in \cite{Blackburn} below.
\subsection{Tree-like property}
A tree-like model is a model whose underlying frame is a tree. If $H$ is a tree with root $r$, we write \HOLinline{\HOLConst{tree}\;\HOLFreeVar{H}\;\HOLFreeVar{r}} in HOL:
\begin{defn}
{\upshape\cite[Definition 1.7]{Blackburn}}
\begin{holmath}
   \HOLConst{tree}\;\HOLFreeVar{H}\;\HOLFreeVar{r}\;\HOLTokenDefEquality{}\\
\;\;\HOLFreeVar{r}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenImp{}}\;{\HOLFreeVar{H}.\HOLFieldName{rel}}\ensuremath{\mid}\sb{\HOLFreeVar{H}.\HOLFieldName{world}}\HOLSymConst{\HOLTokenSupStar{}}\;\HOLFreeVar{r}\;\HOLBoundVar{w})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenNeg{}}\HOLFreeVar{H}.\HOLFieldName{rel}\;\HOLBoundVar{w}\;\HOLFreeVar{r})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{w}.\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{w}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLFreeVar{r}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenUnique{}}\HOLBoundVar{w}\sb{\mathrm{0}}.\;\HOLBoundVar{w}\sb{\mathrm{0}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{H}.\HOLFieldName{world}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{H}.\HOLFieldName{rel}\;\HOLBoundVar{w}\sb{\mathrm{0}}\;\HOLBoundVar{w}
\end{holmath}
\end{defn}
The tree-like property says an satisfiable modal formula can be satisfied in a tree-like model. It is proved in HOL as:
\begin{prop}
{\upshape\cite[Proposition 2.15]{Blackburn}}\label{2.15}
\begin{holmath}
  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :\ensuremath{\beta}\;\HOLTyOp{model}),(\HOLFreeVar{w} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :\ensuremath{\beta}\;\HOLTyOp{list}\;\HOLTyOp{model})\;(\HOLBoundVar{r} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{tree}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}.\HOLFieldName{frame}\;\HOLBoundVar{r}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{r}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
\end{holmath}
\end{prop}
The world set of the tree like model constructed from \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} according to the standard proof of this theorem is a set of lists of worlds in \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}}, that is why passing to a tree-like model does not preserve the type.

@@

\subsection{Finite model property}
There are two classical approach of constructing finite models using model theory, namely via selection and via filtration. 
\subsubsection{Selection}
Given \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\sb{\mathrm{1}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}, where $\phi$ has degree $k$, we can construct \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}}, \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}}, \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}} and \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} comsecutively, such that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} is the finite model we want, where:
\begin{itemize}
\item \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} is the tree-like model obtained from \ref{2.15} with root \HOLinline{\HOLFreeVar{w}\sb{\mathrm{2}}} such that \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLFreeVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}.
\item \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}} is the restriction of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} to height $k$.
\item \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{4}}} is obtained from \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}} by modifying the valuation into \HOLinline{\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{v}.\;\HOLKeyword{if}\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLConst{prop_letters}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLKeyword{then}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sb{\mathrm{3}}}}\HOLSymConst{\ensuremath{\sp{V}}}\;\HOLBoundVar{p}\;\HOLBoundVar{v}\;\HOLKeyword{else}\;\HOLConst{F}}.
\end{itemize}
The construction of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{5}}} requires a lemma: 
\begin{lm}
{\upshape\cite[Proposition 2.29]{Blackburn}}
\begin{holmath}
  \ensuremath{\HOLTokenTurnstile}\HOLConst{FINITE}\;(\HOLFreeVar{\ensuremath{\Phi}} :\HOLTyOp{num}\;\ensuremath{\rightarrow}\;\HOLTyOp{bool})\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{INFINITE}\;\ensuremath{{\cal{U}}}({:}\ensuremath{\beta})\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}(\HOLBoundVar{n} :\HOLTyOp{num}).\\
\;\;\;\;\;\;\;\;\;\;\;\HOLConst{FINITE}\;(\HOLTokenLeftbrace{}\HOLBoundVar{\ensuremath{\phi}}\;\HOLTokenBar{}\;\HOLConst{DEG}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenLeq{}}\;\HOLBoundVar{n}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{prop_letters}\;\HOLBoundVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenSubset{}}\;\HOLFreeVar{\ensuremath{\Phi}}\HOLTokenRightbrace{}\HOLSymConst{//E}\;(\HOLFreeVar{\ensuremath{\mu}} :\ensuremath{\beta}\;\HOLTyOp{itself}))
\end{holmath}
\end{lm}
We require the assumption that the universe of $\beta$ is infinite since we relies on the fact that two modal formulas \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{1}}}} and \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{2}}}} are equivalent if and only if \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{1}}}} and \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{2}}}} are equivalent. This is easy to prove using set theory. However, in simple type theory, the proof of  \HOLinline{\HOLFreeVar{\ensuremath{\phi\sb{1}}}\equiv\sb{({:}\ensuremath{\beta})}\HOLFreeVar{\ensuremath{\phi\sb{2}}}} iff \HOLinline{\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{1}}}\equiv\sb{({:}\ensuremath{\beta})}\HOLSymConst{\ensuremath{\Diamond}}\HOLFreeVar{\ensuremath{\phi\sb{2}}}} requires the infiniteness of the type universe of \HOLinline{\ensuremath{\beta}}.

@@

As we used \ref{2.15}, we change the type of the model by passing to a finite model via selection:
\begin{thm}
{\upshape\cite[Theorem 2.34 (Finite model property, via selection)]{Blackburn}}
\begin{holmath}
  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :\ensuremath{\beta}\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}} :\ensuremath{\beta}\;\HOLTyOp{list}\;\HOLTyOp{model})\;(\HOLBoundVar{v} :\ensuremath{\beta}\;\HOLTyOp{list}).\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{v}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}},\HOLBoundVar{v}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}
\end{holmath}
\end{thm}

\subsubsection{Filtration}
Given \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}, the finite model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} constructed by selection is the filtration of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}} by the set of subformulas of \HOLinline{\HOLFreeVar{\ensuremath{\phi}}}. In HOL, we write \HOLinline{\HOLConst{FLT}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{\ensuremath{\Sigma}}} for the filtration of the model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} via the set \HOLinline{\HOLFreeVar{\ensuremath{\Sigma}}}. The formalization is almost a direct translation of the mathematical proof in Blackburn \emph{et al}~\cite{Blackburn}. For the only difference: the world set of the filtrated model using the standard proof is a set of equivalence classes of worlds. To make the type of the finite model same as the model we start with, instead of taking the set of equivalence classes as the world set, we pick only one representative from each equivalence class. Hence we can prove:
\begin{thm}
{\upshape\cite[Theorem 2.41 (Finite model property, via filtration)]{Blackburn}}
\begin{holmath}
  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :\ensuremath{\beta}\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}} :\ensuremath{\beta}\;\HOLTyOp{model})\;(\HOLBoundVar{w}\sb{\mathrm{2}} :\ensuremath{\beta}).\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLBoundVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}
\end{holmath}
\end{thm}
where \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} above is taken as \HOLinline{\HOLConst{FLT}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\;(\HOLConst{subforms}\;\HOLFreeVar{\ensuremath{\phi}})}, and has the same type as \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}}.
